Re-examining cross-cultural similarity judgements using lexical co-occurrence

### 2023-01-20 code/4_paper/manuscript/manuscript.tex
cogsci paper exchange

\usepackage[utf8]{inputenc}
\usepackage[T1, T5]{fontenc}

@@

\begin{abstract}
Is ``cow'' more closely related to ``grass'' or ``chicken''? Speakers of
different languages judge similarity in this context differently, but
why? One possibility is that cultures co-varying with these languages
induce variation in conceptualizations of similarity. Specifically, East
Asian cultures may promote reasoning about thematic similarity, by which
cow and grass are more related whereas Western cultures may bias
similarity judgements toward taxonomic relations, like cow-chicken. This
difference in notions of similarity is the consensus interpretation for
cross-cultural variation in this paradigm. We consider, and provide
evidence for, an alternative possibility, by which notions of similarity
are equivalent across contexts, but the statistics of the environment
vary. On this account, similarity judgements are guided by co-occurrence
in experience, and observing or hearing about cows and grass or cows and
chickens more often could induce preferences for the relevant grouping,
and account for apparent differences in notions of similarity across
contexts.

@@

Taxonomic and thematic similarity provide a convenient entry point to
broader debates about cross-linguistic and cross-cultural variation in
notions of similarity. Taxonomic categorization is based on the
similarity of attributes, for example, similar perceptual properties,
like shared color or shape, among objects. In contrast, thematic
categorization is based on causal, spatial, and temporal relationships
among objects (Markman \& Hutchinson, 1984). To illustrate this point,
if a participant is given the words ``chicken'' and ``grass'', and asked
which of those words are more closely related to the word ``cow'', the
pairing ``chicken'' - ``cow'' indicates a taxonomic match, whereas
``grass'' - ``cow'' indicates a thematic match.

Ji, Zhang, \& Nisbett (2004) found that preferences for taxonomic and
thematic matches differ across cultures, even when controlling for
testing language and fluency. Specifically, Chinese participants made
more thematic matches than taxonomic, while the converse was true for
European Americans. This difference held even for Chinese participants
tested in English (rather than Mandarin), suggesting that culture has an
effect on the categorization strategy independent of the test language.
However, the study also found an effect of language within Chinese
participants from mainland China (when tested both in China and in the
US): their preference for thematic categorization was significantly
lower when tested in English compared to Mandarin Chinese. This suggests
that test language also influences categorization strategies. Ji et
al.~characterize these results as demonstrating that culture
(independent of language) leads to different styles of reasoning about
similarity, whereas language serves as a ``tuning'' mechanism operating
within a culturally-specific style, by activating representations
corresponding to the language being used.

@@

The view that culture drives differences in categorization style has
been one favored by other non-linguistic studies that use taxonomic vs
thematic categorization to probe similarity judgment. When shown
pictorial triads that represent words (i.e.~pictures of an ant, a bee,
and honey), Chinese children (9-10 years old) scored higher on thematic
categorization compared to American children of the same age range, and
vice versa for taxonomic categorization (Chiu, 1972). These
cross-cultural differences in similarity judgment persist in novel
object categorization, with Chinese participants preferring to group by
family resemblance across multiple features and Americans preferring a
single-feature rule (Norenzayan, Smith, Kim, \& Nisbett, 2002). Since
differences in categorization style persist with novel stimuli (as
opposed to previously seen words or pictures of concepts), this suggests
that participants from different cultural backgrounds are utilizing
different reasoning strategies when making similarity judgments. The
authors attribute these differences to tendencies toward analytic
processing in Western cultures, which emphasizes objects and their
properties and holistic processing in East Asian cultures, which
emphasizes relationships between objects and their context. According to
this framework, East Asians attend more to relationships between objects
and context and reason more holistically, whereas Westerners attend more
to objects and their properties, and reason analytically (ibid, see also
Nisbett (2003)). In related work, East Asian participants show a higher
level of sensitivity to context than their Western counterparts when
reproducing drawings from memory (Ji, Peng, \& Nisbett, 2000); visually
exploring naturalistic scenes (Chua, Boland, \& Nisbett, 2005);
describing scenes (Masuda \& Nisbett, 2001); and in explaining the
causes of ambiguous behaviors (Choi, Nisbett, \& Norenzayan, 1999).
These cross-cultural differences in attention and analytic vs holistic
processing are seen to map onto the cross-cultural differences in
similarity judgment: because East Asian participants attend more to the
context between objects and reason more holistically, it stands to
reason that they would consider words that appear in the same context
(i.e.~thematically similar) to be more similar than
taxonomically-similar words.

This view ascribes cross-cultural differences in similarity judgment
tasks to variation in the conceptualization of similarity itself.
Alternatively, these judgments could be shaped by cross-cultural
differences in the statistics of the environment, and accordingly the
content of everyday experiences. Perhaps when confronted with the triad
task, participants from all cultures follow the same process for
reasoning about similarity, but rely on language or culture-specific
inputs to this process. For example, it may be that Vietnamese
participants are more likely to have knowledge of farming practices than
Americans, and may associate cows and grass more strongly (as grass is
the typical feed for cows in the small-scale family farming contexts
common in Vietnam). Americans, on the other hand, may have strong
associations between cows and chickens as a result of encountering them
together in children's books or petting zoo contexts. If we observe a
difference in categorization between, say Vietnamese and US
participants, it could be that members of both groups use the same
procedure (considering similarity that is influenced by both taxonomic
and thematic relations), but the inputs to this procedure differ between
cultures, with Vietnamese participants exposed to more instances of
``cow'' and ``grass'' occurring together, compared to US participants.

\hypertarget{language-as-a-proxy-of-variation-in-experience}{%
\subsection{Language as a proxy of variation in
experience}\label{language-as-a-proxy-of-variation-in-experience}}

While co-occurrences in experience are difficult to measure,
co-occurrence in language can provide a rough proxy--and indeed,
language may afford many of the ``experiences'' that people have with
infrequently encountered items, like cows or helicopters. In this study,
we use co-occurrence in language as a proxy of cultural experience. This
is a generalizing assumption, and while no linguistic corpus can be
expected to fully capture a culture, using this proxy as a model of
culture provides a conservative test of the hypothesis that differences
in the input, rather than the process, of reasoning produces cultural
variation in this task. This means that if such a model does work -- if
linguistic co-occurrence can predict cultural variation in this
similarity task without building differences in the decision process
into the model -- it is relatively strong evidence that language alone
(perhaps as a more easily observed source of information on broader
cultural and ecological variation) may account for these cross-cultural
differences in reasoning.\footnote{While we discuss cross-cultural
  variability at the level of countries or larger world areas, these are
  not cultural monoliths. For convenience, we operationalize culture at
  the level of country, based on where participants were raised. It is
  an open question whether performance in our participant populations
  (of relatively young and well-educated adults) is representative of
  the broader country. This is especially true for societies with
  substantial ethnic and cultural variation such as the US. We expect
  that our data is likely to underestimate variation both within and
  between the countries we sample from.}

Indeed, previous work suggests that using lexical co-occurrence as a
proxy can be useful in thinking about similarity judgments. Natural
language processing tasks have found that lexical co-occurrence is a
good predictor for performance in similarity judgment tasks. Griffiths,
Steyvers, \& Tenenbaum (2007) showed that a model that takes
word-document co-occurrence as input can be used to predict word
association and the effects of semantic association on a variety of
linguistic tasks. Additionally, lexical semantic models that are
constructed using lexical co-occurrence (as opposed to annotated
relations) have been shown to perform well on predicting human judgments
about similarity between word pairs that are either thematically or
taxonomically related (Rohde, Gonnerman, \& Plaut, 2006).

Our current study will be using 2 measures of lexical co-occurrence: 1)
raw lexical co-occurrences and 2) cosined distances of fastText word
vectors. fastText is a system that uses lexical co-occurrence
information to generate a vector representing each word in its lexicon
(Mikolov, Grave, Bojanowski, Puhrsch, \& Joulin, 2018). fastText uses a
continuous bag-of-words model, which takes into account a symmetric
window of words surrounding the word in question and maximizes the
log-likelihood of the probability of the word given its context.
Calculating the cosine-similarity between word vectors from fastText
gives us a score of how similar the contexts of these words are.
Therefore, words that occur in the same context (high lexical
co-occurrence) would have a larger cosine-similarity score.

It must be noted that fastText supplements the lexical co-occurrence
information by simultaneously training weight vectors for each position
relative to the word in question, as well as adding sub-word information
(bag-of-character n-gram vectors for each character forming the words).
The similarity judgment prediction given by fastText may be more higher
than a model using purely lexical co-occurrence information. However,
models that use only lexical co-occurrence information (with or without
a simple position weight system that gives more weight to words that are
closer to the word in question) also perform well in similarity
prediction tasks. For this study, due to the lack of large cleaned
corpora for Mandarin and Vietnamese, we are leveraging fastText
pre-trained models on the languages we are interested in Grave,
Bojanowski, Gupta, Joulin, \& Mikolov (2018).

@@

Ji et al. (2004) suggest that the notion of similarity is what varies
across cultural groups, but it is quite possible that similarity is
regarded in the same way and it is actually the input to this similarity
judgment that varies across culture contexts. It might be possible to
gain traction on possible mechanisms by examining whether variation in
similarity judgments co-varies with environmental statistics that differ
across the cultural and linguistic contexts.

@@

We then ask to what extent similarity judgments show cross-context
differences that align with co-occurrence patterns in participants'
languages. We would only see a main effect of cultural context if
differing notions of similarity drive these cross-cultural differences.
Meanwhile, if we do see variation in responding trial-to-trial that
tracks with lexical co-occurrence statistics (as a rough proxy combining
both linguistic and cultural context), this suggests that cross-cultural
notions of similarity might be similar, and it is environmental
statistics that guide the differences in similarity judgments. To
preview our results, we find that, while we do not see overall group
differences (between the East Asian contexts and the US) in similarity,
we do find language-specific co-occurrence can indeed explain
cross-context differences in similarity judgments. In contrast to
previous accounts, by which culture induces differing conceptions of
similarity (varying between taxonomic and thematic), these findings
provide an alternative, and more specific, account by which language may
explain these cross-cultural differences without invoking variation in
notions of similarity.

@@

We excluded 8 US participants, 62 Vietnam participants, and 34 China
participants who did not answer all attention checks correctly. We
followed 4 exclusion criteria that aim to retain only participants who
are influenced by one culture: (1) non-native speakers of English and
Vietnamese, respectively, (2) fluent in at least one of the other two
study languages (Vietnamese for US participants, English for Vietnamese
participants and Chinese participants), (3) have lived outside of the
test country (US, Vietnam, or China) for more than two years, and (4)
have significant international experience (more than 6 international
experiences of 2 days or longer.) We did not use a particular criterion
for a language if it would exclude 25\% or more of any one sample. In
this round of exclusion, we excluded 68 US, 27 Vietnam participants, and
34 China participants. After these exclusions, the US sample included
124 participants (30M, 89F, 3 non-binary, 2 other), with mean age =
22.18 (SD = 8) and median age = 20. The Vietnam sample included 110
participants (34M, 71F, 5 other), with mean age = 22.21 (SD = 5.81) and
median age = 21. The China sample included 132 participants (51M, 81F,
other), with mean age = 23.27 (SD = 3.75) and median age = 23.

Notably, we lost a majority of our Vietnam participants (more than 60\%)
in the attention check exclusion. One reason we suspect why this might
have happened is because our Vietnam participants are less familiar with
research surveys and attention check questions, and thus might have
thought too much about the attention check questions. We carried out an
exploratory analysis where we used a less stringent attention check
exclusion (covered in the final part of the Results \& Analysis
section). To preview our findings for this analysis, we did not find any
difference in significant results when compared with the main analysis.

@@

We adapted stimuli from previous studies to create a set of test triads
consisting of a cue, with one thematic and one taxonomic match option.
For example, ``cow,'' ``grass,'' and ``chicken,'' where ``cow'' is the
cue, ``grass'' is the thematic match, and ``chicken'' the taxonomic
match. We included 105 such triads, a superset including triads pulled
from supplemental information and in-text examples across the
literature, and others that we adapted or created (see SI for the full
list of stimuli and sources). We selected triads on the basis of
cultural familiarity in the US, Vietnam, and China. The triads were
originally in English; they were translated to Vietnamese and Mandarin
by a fluent bilingual speaker in each language. The translations were
then checked for accuracy after backtranslation to English by another
fluent bilingual in each language who was naive to the original English
versions.

@@

Our general approach is to build a model of similarity that is based on
collocation counts in each language.

\hypertarget{raw-lexical-co-occurrences}{%
\subsubsection{Raw lexical
co-occurrences}\label{raw-lexical-co-occurrences}}

To give an intuition for our model, consider the cow-grass-chicken
triad: we counted how many times ``cow'' and ``grass'' co-occur within a
window of text in each corpus, and compare this to how many times
``cow'' and ``chicken'' co-occur. Our similarity prediction is then
proportional to the relative frequency of these pairs. For example, if
the thematic cow-grass match accounts for 30\% of collocations for this
triad (with cow-chicken making up the other 70\%), then our model
predicts, correspondingly, that 30\% of responses to the triad will be
grass, and the other 70\% chicken. We then use a mixed-effects
regression to evaluate how well each corpus collocate model predicts
participants' similarity judgments, across triads and languages.

\hypertarget{collocate-retrieval-and-coding}{%
\paragraph{Collocate retrieval and
coding}\label{collocate-retrieval-and-coding}}

For our English co-occurrence metric, we used the online interface of
the Corpus of Contemporary American English (Davies, 2008-) to retrieve
collocation counts. We recorded the raw count of times that any
cue-match pair (e.g., cow-grass or cow-chicken) co-occur in a window of
19 words, which is the maximum window size in the online interface, and
closest to the sentence co-occurrence metric in our VI corpus
collocation counts.

To determine Vietnamese co-occurrence, we used the raw frequency of
sentence co-occurrences from a subset of the Vietnamese corpus in the
Leipzig Corpora Collection (Goldhahn, Eckart, \& Quasthoff, 2012); this
corpus includes 70 million Vietnamese sentences, but our corpus data
comes from a 1 million sentence subset for which co-occurrence counts
are available for download. Vietnamese makes very frequent use of
compositional morphology but the written language uses spacing to
delineate syllable boundaries rather than word boundaries. Accordingly,
collocate searches returned instances of both target terms and many
morphologically related, but distinct, words. We included in our counts
any instances of the target term or close semantic neighbors containing
the same morpheme(s) as long as they entailed a likely literal reference
to the target term. For example, our search for collocates of the term
``gà'' (chicken) also returned ``gà mái,'' a distinct word for female
chickens. Despite being a different word, ``gà mái'' both includes the
morpheme for ``chicken'' and entails reference to a chicken.
Accordingly, instances of both ``gà'' and ``gà mái'' were included in
our collocate count for ``chicken.'' Some compounds do not meet this
criteria, and were excluded from collocate counts. For example ``trái
cây'' (tree.fruits) includes the same syllable as the word ``cây''
(tree), but refers to fruit that comes from trees, not the trees
themselves, and was therefore excluded from our collocate counts for
``tree.''

TODO; ADD INFO ABOUT MANDARIN CORPUS

\hypertarget{collocate-similarity-model}{%
\paragraph{Collocate similarity
model}\label{collocate-similarity-model}}

From the raw co-occurrence counts of each triad, we calculated the
thematic co-occurrence ratio as the number of thematic co-occurrences
over the sum of thematic and taxonomic co-occurrences. We did this for
both the English and Vietnamese corpus co-occurrence counts. In this
way, we obtained predictions for 73 of 105 triads from the Vietnamese
corpus and 104 of 105 triads from English. We therefore limited all
analyses reported here to the subset of triads for which our
corpus-based model can make meaningful predictions, meaning triads that
have at least one non-zero collocate count (either thematic or taxonomic
match). We replaced any remaining zero collocate counts with
\(\epsilon\) to account for sparsity in the corpus data. We then tested
whether these simple relative frequency models predict US and VN
responding.

\hypertarget{cosine-distance-of-word-vectors}{%
\subsubsection{Cosine distance of word
vectors}\label{cosine-distance-of-word-vectors}}

@@

\hypertarget{collocate-retrieval}{%
\paragraph{Collocate retrieval}\label{collocate-retrieval}}

We use the fastText pre-trained models of English, Mandarin, and
Vietnamese in Grave et al. (2018). These models are trained on Common
Crawl and Wikipedia using We distribute pre-trained word vectors for 157
languages, trained on Common Crawl and Wikipedia using fastText. These
models were trained using a Continuous Bag of Words (CBOW) with
position-weights and a window of size 5. The models use character
n-grams of length 5 and 10 negative examples.From the aforementioned
models, we retrieve the word vectors (dimension 300) for each word we
are interested in.

\hypertarget{collocate-similarity-model-1}{%
\paragraph{Collocate similarity
model}\label{collocate-similarity-model-1}}

@@

The group means of proportion of thematic response in mainland China is
the highest (M = 0.65, SD = 0.12), followed by the groups means in
Vietnam (M = 0.6, SD = 0.11), which is slightly higher than that of the
US (M = 0.57, SD = 0.16).

\begin{CodeChunk}
\begin{figure}[tb]
\includegraphics{figs/unnamed-chunk-2-1} \caption[Proportion of thematic responses by country]{Proportion of thematic responses by country. Only US-China responding comparison shows a siginficant difference. We could not extend this to US-Vietnam responding comparison.}\label{fig:unnamed-chunk-2}
\end{figure}
\end{CodeChunk}

@@

Overall, there is a significant effect of country on proportion of
thematic responses (\(\chi^2\)(2) = 15.75, p \textless{} .001). However,
this effect is driven by the difference between US and China responding
(\(\beta\) = -0.5, p \textless{} .001). There is no statistical
difference between the Vietnam and China responding (\(\beta\) = -0.25,
p = 0.066), and the US and Vietnam responding (\(\beta\) = 0.25, p =
0.095).

@@

\hypertarget{is-responding-in-each-cultural-context-predicted-by-the-corresponding-lexical-statistics-raw-co-occurrences}{%
\subsubsection{Is responding in each cultural context predicted by the
corresponding lexical statistics (raw
co-occurrences)?}\label{is-responding-in-each-cultural-context-predicted-by-the-corresponding-lexical-statistics-raw-co-occurrences}}

\hypertarget{raw-lexical-co-occurrences-1}{%
\paragraph{Raw lexical
co-occurrences}\label{raw-lexical-co-occurrences-1}}

To test whether variation in lexical co-occurrence can explain
differences in similarity judgments between US and Vietnam participants,
we compare logistic mixed-effects regression models fit to the thematic
responding data from each country separately. We first ask how well each
corpus model (English, Vietnamese, or Mandarin) predicts similarity
judgments by speakers of the corresponding language (US, Vietnam, or
China). To do this, we use a mixed-effects logistic regression to
predict triad responses (0=taxonomic or 1=thematic) with corpus
prediction (proportion thematic co-occurrence or cosine distance) as a
fixed effect and participant and triad as random effects.

We found that all corpora are significant predictors of all cultural
context responding.

For US responding: English (EN), Vietnamese (VI), and Mandarin (ZH)
corpus are all significant predictors. EN corpus: \(\beta\) = 2.01,
\(\chi^2\)(1) = 17.42, p \textless{} .001. VI corpus: \(\beta\) = 1.21,
\(\chi^2\)(1) = 11.1, p \textless{} .001. ZH corpus: \(\beta\) = 1.15,
\(\chi^2\)(1) = 6.38, p = 0.012.

For Vietnam responding: English (EN), Vietnamese (VI), and Mandarin (ZH)
corpus are all significant predictors. EN corpus: \(\beta\) = 1.73,
\(\chi^2\)(1) = 12.35, p \textless{} .001. VI corpus: \(\beta\) = 1.27,
\(\chi^2\)(1) = 12.83, p \textless{} .001. ZH corpus: \(\beta\) = 1.8,
\(\chi^2\)(1) = 18, p \textless{} .001.

For China responding: English (EN), Vietnamese (VI), and Mandarin (ZH)
corpus are all significant predictors. EN corpus: \(\beta\) = 1.21,
\(\chi^2\)(1) = 9.83, p = 0.002. VI corpus: \(\beta\) = 0.77,
\(\chi^2\)(1) = 7.44, p = 0.006. ZH corpus: \(\beta\) = 1.15,
\(\chi^2\)(1) = 13.28, p \textless{} .001.

This suggests a strong shared similarity signal across the languages.

\hypertarget{cosine-distances-of-fasttext-word-vectors}{%
\paragraph{Cosine distances of fastText word
vectors}\label{cosine-distances-of-fasttext-word-vectors}}

Similar to the results from the analysis with raw lexical
co-occurrences, we found that all corpora are significant predictors of
all cultural context responding.

For US responding: English (EN), Vietnamese (VI), and Mandarin (ZH)
corpus are significant predictors of US responding (EN-US model:
\(\beta\) = -8.64, \(\chi^2\)(1) = 30.44, p \textless{} .001; VI-US
model: \(\beta\) = -2.97, \(\chi^2\)(1) = 5.23, p = = 0.022; ZH-US
model: \(\beta\) = -6.96, \(\chi^2\)(1) = 18.08, p \textless{} .001).

For Vietnam (VN) responding: English (EN), Vietnamese (VI), and Mandarin
(ZH) corpus are significant predictors of VN responding (EN-VN model:
\(\beta\) = -5.33, \(\chi^2\)(1) = 9.39, p = 0.002; VI-VN model:
\(\beta\) = -3.46, \(\chi^2\)(1) = 6.64, p = = 0.01; ZH-VN model:
\(\beta\) = -6.37, \(\chi^2\)(1) = 13.93, p \textless{} .001).

For China (CN) responding: English (EN), Vietnamese (VI), and Mandarin
(ZH) corpus are significant predictors of CN responding (EN-CN model:
\(\beta\) = -6.18, \(\chi^2\)(1) = 25.34, p \textless{} .001; VI-CN
model: \(\beta\) = -2.44, \(\chi^2\)(1) = 5.83, p = = 0.016; ZH-CN
model: \(\beta\) = -7.36, \(\chi^2\)(1) = 37.45, p \textless{} .001).

\hypertarget{is-responding-in-each-cultural-context-best-predicted-by-the-corresponding-corpuss-lexical-statistics-raw-co-occurrences-as-opposed-to-the-other-two-corpora}{%
\subsubsection{\texorpdfstring{Is responding in each cultural context
\textbf{best} predicted by the corresponding corpus's lexical statistics
(raw co-occurrences), as opposed to the other two
corpora?}{Is responding in each cultural context best predicted by the corresponding corpus's lexical statistics (raw co-occurrences), as opposed to the other two corpora?}}\label{is-responding-in-each-cultural-context-best-predicted-by-the-corresponding-corpuss-lexical-statistics-raw-co-occurrences-as-opposed-to-the-other-two-corpora}}

@@

>>>>

\usepackage[utf8]{inputenc}
\usepackage[T1, T2A, T5]{fontenc}
\usepackage{kotex}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

@@

\begin{abstract}
Is ``cow'' more closely related to ``grass'' or ``chicken''? Speakers of
different languages judge similarity in this context differently, but
why? One possibility is that cultures co-varying with these languages
induce differences in conceptualizations of similarity. Specifically,
East Asian cultures may promote reasoning about thematic similarity, by
which cow and grass are more related, whereas Western cultures may bias
judgements toward taxonomic relations, like cow-chicken. This difference
in notions of similarity is the consensus interpretation for
cross-cultural variation in this paradigm. We consider, and provide
evidence for, an alternative possibility, by which notions of similarity
are similar across contexts, but the statistics of the environment vary.
On this account, similarity judgements are guided by co-occurrence in
experience, and observing or hearing about cows and grass or cows and
chickens more often could induce preferences for these groupings, and
account in part for apparent differences in notions of similarity across
contexts.

@@

Taxonomic and thematic similarity provide a convenient entry point to
broader debates about cross-linguistic and cross-cultural variation in
notions of similarity. Taxonomic categorization is based on the
similarity of attributes, for example, similar perceptual properties,
like shared color or shape, among objects. In contrast, thematic
categorization is based on causal, spatial, and temporal relationships
among objects (Markman \& Hutchinson, 1984).

@@

Preferences for taxonomic and thematic similarity vary across cultures.
Chiu (1972) found that Chinese children (9-10 years old) are more likely
to choose thematic matches in a picture triad task (shown, e.g., images
of an ant, a bee, and honey) than their American counterparts. These
cross-cultural differences are also observed in novel object
categorization, with Chinese participants preferring to group by family
resemblance across multiple features and Americans preferring a
single-feature rule (Norenzayan, Smith, Kim, \& Nisbett, 2002). The
authors link these differences to tendencies toward analytic processing
in Western cultures, which emphasizes objects and their properties, and
holistic processing in East Asian cultures, which emphasizes
relationships between objects and their context (see also Nisbett
(2003)). In related work, East Asian participants show a higher level of
sensitivity to context than their Western counterparts when reproducing
drawings from memory (Ji, Peng, \& Nisbett, 2000); visually exploring
naturalistic scenes (Chua, Boland, \& Nisbett, 2005); describing scenes
(Masuda \& Nisbett, 2001); and in explaining the causes of ambiguous
behaviors (Choi, Nisbett, \& Norenzayan, 1999).

Ji, Zhang, \& Nisbett (2004) asked whether differences in analytic and
holistic processing are driven by language, culture, or a combination of
these. They presented Chinese and European American adults with a word
triad task, and found a preference for thematic matching among Chinese
participants compared to European Americans. They found that test
language contributed to this difference, but did not fully explain it:
Chinese participants showed a stronger thematic preference when tested
in Mandarin, but still showed a thematic preference when tested in
English. Ji et al.~conclude that culture (independent of language) leads
to different styles of reasoning about similarity, whereas language
serves as a ``tuning'' mechanism operating within a culturally-specific
style, by activating representations corresponding to the language being
used.\footnote{But note that this work did not include a manipulation to
  prevent participants from engaging in e.g., covert naming in a
  language other than the test language.}

This view ascribes cross-cultural differences in similarity judgment to
variation in the conceptualization of similarity itself. Alternatively,
these judgments could be shaped by cross-cultural differences in the
statistics of the environment, and accordingly the content of everyday
experiences. Perhaps when confronted with the triad task, participants
from all cultures follow the same process for reasoning about
similarity, but rely on language or culture-specific inputs to this
process. If we observe a difference in categorization between East Asian
and Western participants, it could be that members of both groups use
the same procedure (considering similarity that is influenced by both
taxonomic and thematic relations), but the inputs to this procedure
differ between cultures, with East Asian participants exposed to more
instances of thematic similarity than their Western counterparts.

\hypertarget{estimating-variation-in-experience}{%
\subsection{Estimating variation in
experience}\label{estimating-variation-in-experience}}

While co-occurrences in experience are difficult to measure,
co-occurrence in language can provide a rough proxy--and indeed,
language may afford many of the ``experiences'' that people have with
infrequently encountered items, like cows or helicopters. In this study,
we use co-occurrence in language as a proxy of cultural experience. This
is a generalizing assumption, and while no linguistic corpus can be
expected to fully capture a culture, using this proxy as a model of
culture provides a conservative test of the hypothesis that differences
in the input, rather than the process, of reasoning produces cultural
variation in this task. This means that if such a model does work -- if
linguistic co-occurrence can predict cultural variation in this
similarity task without building differences in the decision process
into the model -- it is relatively strong evidence that language alone
(perhaps as a more easily observed source of information on broader
cultural and ecological variation) may account for these cross-cultural
differences in reasoning.

Indeed, previous work suggests that using lexical co-occurrence as a
proxy can be useful in thinking about similarity reasoning. Natural
language processing tasks have found that lexical co-occurrence is a
good predictor for performance in similarity judgment. Griffiths,
Steyvers, \& Tenenbaum (2007) showed that a model that takes
word-document co-occurrence as input can be used to predict word
association and the effects of semantic association on a variety of
linguistic tasks. Additionally, lexical semantic models that are
constructed using lexical co-occurrence (as opposed to annotated
relations) have been shown to perform well on predicting human judgments
about similarity between word pairs that are thematically or
taxonomically related (Rohde, Gonnerman, \& Plaut, 2006).

Our study uses cosine distances of fastText word vectors as a measure of
lexical co-occurrence\footnote{We also carried out our analysis using
  raw lexical co-occurrences and obtained similar results.}. fastText is
a system that uses lexical co-occurrence information to generate a
vector representing each word in its lexicon (Mikolov, Grave,
Bojanowski, Puhrsch, \& Joulin, 2018).

@@

Ji et al. (2004) suggest that the notion of similarity is what varies
across cultural groups, but it is possible that similarity is regarded
in the same way and it is actually the input to this similarity judgment
that varies across cultural contexts. It may be possible to gain
traction on potential mechanisms by examining whether variation in
similarity judgments co-vary with environmental statistics that differ
across cultural and linguistic contexts.

@@

We then ask to what extent similarity judgments show cross-context
differences that align with co-occurrence patterns in participants'
languages. We would only see a main effect of cultural context if
differing notions of similarity drive these cross-cultural differences.
Meanwhile, if we do see variation in responding trial-to-trial that
tracks with lexical co-occurrence statistics (as a rough proxy combining
both linguistic and cultural context), this suggests that cross-cultural
notions of similarity might be similar, and it is environmental
statistics that guide the differences in similarity judgments.

To preview our results, we find that, while we do not see overall group
differences (between the East Asian contexts and the US) in similarity,
we do find language-specific co-occurrence can indeed explain
cross-context differences in similarity judgments. In contrast to
previous accounts, by which culture induces differing conceptions of
similarity (varying between taxonomic and thematic), these findings
provide an alternative, and more specific, account by which language may
explain these cross-cultural differences without invoking variation in
notions of similarity.

There are several important limitations of our approach. While we
discuss cross-cultural variability at the level of countries or larger
world areas, these are not cultural monoliths. For convenience, we
operationalize culture at the level of country, based on where
participants were raised. It is an open question whether performance in
our participant populations (of relatively young and well-educated
adults) is representative of the broader country. This is especially
true for societies with substantial ethnic and cultural variation such
as the US. We expect that our data is likely to underestimate variation
both within and between the countries we sample from.

Language, culture, cognition, and individual experiences are intertwined
in complex causal relationships: cultural features (farming) can lead to
differences in individuals' experiences (seeing cows and grass) and in
turn these can influence language (talking more about cows and grass).
But cultural features (a thematic orientation) could also more directly
cause individuals to talk differently about the same experiences
(mentioning what cows eat rather than what other animals cows are like).
In this study, we measure language and its relation to cross-cultural
differences in categorization, but these relations test only the
plausibility of a language-based account; they cannot establish the
direction of causality.

@@

We excluded 26 US participants, 123 Vietnam participants, and 34 China
participants who did not answer all attention checks correctly. We
followed 4 exclusion criteria that aim to retain only participants who
are influenced by one culture: (1) non-native speakers of English and
Vietnamese, respectively, (2) fluent in at least one of the other two
study languages (Vietnamese for US participants, English for Vietnamese
participants and Chinese participants), (3) have lived outside of the
test country (US, Vietnam, or China) for more than two years, and (4)
have significant international experience (more than 6 international
experiences of 2 days or longer.) We did not use a particular criterion
for a language if it would exclude 25\% or more of any one sample. In
this round of exclusion, we excluded 65 US, 19 Vietnam participants, and
34 China participants. After these exclusions, the US sample included
109 participants (27M, 78F, 3 non-binary, 1 other), with mean age = 22.4
(SD = 8.49) and median age = 20. The Vietnam sample included 57
participants (17M, 40F, other), with mean age = 22.21 (SD = 6.12) and
median age = 21. The China sample included 132 participants (51M, 81F,
other), with mean age = 23.27 (SD = 3.75) and median age =
23.\footnote{A table summarizing number of participants lost at each
  round of exclusions is included in the Appendix.}

Notably, we lost a majority of our Vietnam participants (more than 60\%)
in the attention check exclusion. One reason we suspect why this might
have happened is because our Vietnam participants are less familiar with
research surveys and attention check questions, and thus might have
thought too much about the attention check questions.\footnote{We
  carried out an exploratory analysis where we used a less stringent
  attention check exclusion (covered in the final part of the Results \&
  Analysis section). To preview our findings for this analysis, we did
  not find any difference in significant results when compared with the
  main analysis.}

@@

We adapted stimuli from previous studies to create a set of test triads
consisting of a cue, with one thematic and one taxonomic match option.
For example, ``cow,'' ``grass,'' and ``chicken,'' where ``cow'' is the
cue, ``grass'' is the thematic match, and ``chicken'' the taxonomic
match. We included 105 such triads, a superset including triads pulled
from supplemental information and in-text examples across the
literature, and others that we adapted or created . We selected triads
on the basis of cultural familiarity in the US, Vietnam, and China. The
triads were originally in English; they were translated to Vietnamese
and Mandarin by a fluent bilingual speaker in each language. The
translations were then checked for accuracy after backtranslation to
English by another fluent bilingual in each language who was naive to
the original English versions. --\textgreater{}

@@

@@

\hypertarget{word-vector-retrieval}{%
\paragraph{Word vector retrieval}\label{word-vector-retrieval}}

We use the fastText pre-trained models of English, Mandarin, and
Vietnamese in Grave, Bojanowski, Gupta, Joulin, \& Mikolov (2018). These
models are trained on Common Crawl and Wikipedia using We distribute
pre-trained word vectors for 157 languages, trained on Common Crawl and
Wikipedia using fastText. These models were trained using a Continuous
Bag of Words (CBOW) with position-weights and a window of size 5. The
models use character n-grams of length 5 and 10 negative examples.From
the aforementioned models, we retrieve the word vectors (dimension 300)
for each word we are interested in.

\hypertarget{similarity-model}{%
\paragraph{Similarity model}\label{similarity-model}}

@@

The group means of proportion of thematic response in mainland China is
the highest (M = 0.65, SD = 0.12), followed by the groups means in
Vietnam (M = 0.6, SD = 0.13), which is slightly higher than that of the
US (M = 0.56, SD = 0.17) (Figure 1).

\begin{CodeChunk}
\begin{figure}[tb]

{\centering \includegraphics{figs/unnamed-chunk-1-1}

\caption[Proportion of thematic responses by country]{Proportion of thematic responses by country.}\label{fig:unnamed-chunk-1}
\end{figure}
\end{CodeChunk}

@@

Overall, there is a significant effect of country on proportion of
thematic responses (\(\chi^2\)(2) = 15.22, p \textless{} .001). However,
this effect is driven by the difference between US and China responding
(\(\beta\) = -0.53, p \textless{} .001). There is no statistical
difference between the Vietnam and China responding (\(\beta\) = -0.25,
p = 0.137), and the US and Vietnam responding (\(\beta\) = 0.28, p =
0.142).

@@

\hypertarget{is-responding-in-each-cultural-context-predicted-by-lexical-statistics}{%
\subsubsection{Is responding in each cultural context predicted by
lexical
statistics?}\label{is-responding-in-each-cultural-context-predicted-by-lexical-statistics}}

To test whether variation in lexical statistics can explain differences
in similarity judgments between US and Vietnam participants, we compare
logistic mixed-effects regression models fit to the thematic responding
data from each country separately. We first ask how well each corpus
model (English, Vietnamese, or Mandarin) predicts similarity judgments
by speakers of the corresponding language (US, Vietnam, or China). To do
this, we use a mixed-effects logistic regression to predict triad
responses (0=taxonomic or 1=thematic) with corpus prediction (proportion
of cosine distance) as a fixed effect and participant and triad as
random effects.

We found that all corpora are significant predictors of all cultural
context responding, with p\textless0.05 and \(\beta\) from -9.05 to
-2.44. (For a full report, see Appendix.)

\hypertarget{is-responding-in-each-cultural-context-best-predicted-by-the-corresponding-corpuss-lexical-statistics-as-opposed-to-the-other-two-corpora}{%
\subsubsection{Is responding in each cultural context best predicted by
the corresponding corpus's lexical statistics, as opposed to the other
two
corpora?}\label{is-responding-in-each-cultural-context-best-predicted-by-the-corresponding-corpuss-lexical-statistics-as-opposed-to-the-other-two-corpora}}

@@

\hypertarget{appendix}{%
\section{Appendix}\label{appendix}}
